A rough of the thesis structure is presented below. The bullets under the subsections are general ideas of what should be included in each chapter, not subsubsections.

\begin{itemize}
\item Abstract
\item Preface
\item Contents
\item Abbreviations
\end{itemize}

\begin{enumerate}
\item Introduction
  \subitem Research Problem
  \begin{itemize}[leftmargin=45px]
    \item Distributing stream computing, how to virtualize dataplane applications using fastpath hardware?
    \item Several methods to understand dataplane applications: system inspection, modeling, simulation, measurements
    \item Hard problem, general concepts and own contributions are not in their own sections
  \end{itemize}
  \subitem Contributions
  \subitem Structure

\item Background
  \subitem Parallel Computing
  \begin{itemize}[leftmargin=45px]
    \item Free lunch is over. Why do we parallelize computing in the first place?
    \item MPSoC devices
    \item From threads to tasks. Why do we need new parallel programming methods such as OpenEM?
  \end{itemize}
  \subitem Stream Computing
  \begin{itemize}[leftmargin=45px]
    \item Internet of Things, fog computing.
    \item From Von Neumann to Streams. Window-type view of the data. Real-time.
  \end{itemize}
  \subitem Distributed Computing
  \begin{itemize}[leftmargin=45px]
    \item Distributed computing is widely used e.g. in scientific computing
  \end{itemize}

\item Distributed, Parallel Stream Computing
  \subitem Dataplane
  \begin{itemize}[leftmargin=45px]
    \item Protocols, addresses, ports, packet dropping, flow reordering ...
    \item Reference implementations: Intel DPDK, ODP
  \end{itemize}
  \subitem Fastpath Hardware
  \begin{itemize}[leftmargin=45px]
    \item What is required from the hardware for the distributed computing to be feasible
  \end{itemize}
  \subitem Open Event-Machine
  \begin{itemize}[leftmargin=45px]
    \item Mapping between fastpath hardware and dataplane application
    \item Event driven parallel programming model
    \item Currently no support for distributed computation
  \end{itemize}
  \subitem Message Passing
  \begin{itemize}[leftmargin=45px]
    \item What is actually needed to move the computation between the nodes?
  \end{itemize}

\item Simulation Model
  \subitem PSE - Performance Simulation Environment
  \subitem Reference values
    \subsubitem Measurement Hardware
    \begin{itemize}[leftmargin=45px]
      \item Example of relevant fastpath hardware
      \item NOT an example of balanced distributed (virtualized) fastpath system
      \item But helps us figuring out the relevant values of fastpath system needed for the model
    \end{itemize}
    \subsubitem Communication Latencies
    \begin{itemize}[leftmargin=45px]
      \item Statistical behaviour of input and output under varying workload
      \item Simple tests with message passing. MPI or future-promise
    \end{itemize}
    \subsubitem Memory Latencies
  \subitem Workload model
  \subitem Hardware model
  \begin{itemize}[leftmargin=45px]
    \item Scheduler, input/output, NPU nodes
  \end{itemize}
  \subitem Software model
  \begin{itemize}[leftmargin=45px]
    \item Scheduler, OpenEM-type application model
  \end{itemize}

\item Model Evaluation
  \subitem Simulation Results
  \subitem Discoveries
    \begin{itemize}[leftmargin=45px]
      \item What did we learn from the model about distributing stream computing and virtualizing dataplane?
    \end{itemize}
  \subitem Building prototype system based on the model
    \begin{itemize}[leftmargin=45px]
      \item Discussion about: distribution is unkown in stream computing context -> the model based of reference measurements hopefully shows us how a real prototype system could work.
      \item Show that the components needed to build the model actually exist (as shown earlier in the thesis)
      \item Problems, opportunities...
    \end{itemize}


\item Discussion
  \subitem Challenge
  \subitem Discoveries
  \subitem Related Work
  \subitem Future Work

\item Conclusions
\end{enumerate}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "thesis-plan-hartikainen"
%%% End:
