\chapter{Background}
\label{chapter:background}

\todo[inline]{write introduction to the backgroup chapter}
\todo[inline]{rename the background}

\section{Parallel Computing}
\label{section:parallel-computing}

\todo[inline]{Free lunch is over. Why do we paralellize computing in the first place?}
\todo[inline]{MPSoC devices}
\todo[inline]{From threads to tasks. Why do we need new parallel programming methods such as OpenEM?}

\section{Dataplane}
\label{sec:dataplane}

\todo[inline]{clean API with a set of coherent libraries and drivers}

\todo[inline]{generic support for many CPUs/NICs}
\todo[inline]{easy to use and understand}
\todo[inline]{documentation}

\todo[inline]{DPDK subscribers/commits}

\todo[inline]{Many of today's dataplane architectures use run to completion model. No cycles can be wasted for scheduling. Applications poll for received packets. Power management?}

\todo[inline]{poll mode drivers}

\todo[inline]{DPDK almost ISA neutral, but has to be done properly}

\todo[inline]{DPDK: Hugepages -> less TLB cache misses}

\todo[inline]{Legacy dataplane? Legacy IPsec?}

\todo[inline]{Protocols, addresses, ports, packet dropping, flow reordering}
\todo[inline]{Reference implementations: Intel DPDK, ODP}

\section{Open Event-Machine}

\todo[inline]{vs. Intel DPDK, ODP??}
\todo[inline]{Mapping between fastpath hardware and dataplane application}
\todo[inline]{Event driven parallel programming model}
\todo[inline]{Currently no support for distributed computation}

\section{Stream Computing}

\todo[inline]{Internet of Things, fog computing}
\todo[inline]{From Von Neumann to Streams. Window-type view of the data. Real-time.}

\section{Distributed Computing}

\todo[inline]{Distributed computing is widely used e.g. in scientific computing}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "thesis-hartikainen"
%%% End:
