\chapter{Distributed Parallel Stream Computing}
\label{chapter:distributed-parallel-stream-computing}

\todo[inline]{introduction paragraph for distributed parallel stream computing -section}

\section{Dataplane}
\label{sec:dataplane}

\todo[inline]{clean API with a set of coherent libraries and drivers}

\todo[inline]{generic support for many CPUs/NICs}
\todo[inline]{easy to use and understand}
\todo[inline]{documentation}

\todo[inline]{DPDK subscribers/commits}

\todo[inline]{Many of today's dataplane architectures use run to completion model. No cycles can be wasted for scheduling. Applications poll for received packets. Power management?}

\todo[inline]{poll mode drivers}

\todo[inline]{DPDK almost ISA neutral, but has to be done properly}

\todo[inline]{DPDK: Hugepages -> less TLB cache misses}

\todo[inline]{Legacy dataplane? Legacy IPsec?}

\todo[inline]{Protocols, addresses, ports, packet dropping, flow reordering}
\todo[inline]{Reference implementations: Intel DPDK, ODP}

\section{Fastpath hardware}

\todo[inline]{What is required from the hardware for the distributed computing to be feasible?}

\section{Open Event-Machine}

\todo[inline]{vs. Intel DPDK, ODP??}
\todo[inline]{Mapping between fastpath hardware and dataplane application}
\todo[inline]{Event driven parallel programming model}
\todo[inline]{Currently no support for distributed computation}

\section{Message Passing}

\todo[inline]{What is actually needed to move the computation between the nodes?}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "thesis-hartikainen"
%%% End:
